### 2021-06-09 doc/src/AdvPLIC.tex
Clarify addresses of IMSIC interrupt files, and support major subdivision of address space.
Add recommendations for how to arrange the memory regions of IMSIC
interrupt files in a machine's address space.  Also, modify an
Advanced PLIC's MSI address configuration registers (mmsiaddrcfg,
mmsiaddrcfgh, smsiaddrcfg, smsiaddrcfgh) to support the subdivision
of the machine's address space so that different groups of harts are
constrained to separate parts of the address space (usually because
the harts are on separate chips in a "multi-socket" configuration).

\textbf{%
The parameters used to configure MSI addresses will likely change from
what is currently described here for registers \z{mmsiaddrcfg} and
\z{mmsiaddrcfgh}.%
}
\vspace{3mm}

@@

When implemented, \z{mmsiaddrcfg} has this format:\nopagebreak
\begin{displayLinesTable}[l@{\quad}l]
bits 31:12 & Low Base PPN (\WARL) \\
bits 10:8  & HXS (\WARL) \\
bit 0      & L \\
\end{displayLinesTable}
All other bits of \z{mmsiaddrcfg} are reserved and read as zeros.
Register \z{mmsiaddrcfgh} has this format:\nopagebreak
\begin{displayLinesTable}[l@{\ \quad}l]
bits 31:0 & High Base PPN (\WARL) \\
\end{displayLinesTable}

Fields High Base PPN from \z{mmsiaddrcfgh} and Low Base PPN from
\z{mmsiaddrcfg} concatenate to form a \mbox{52-bit} Base PPN
(Physical Page Number).
The use of this value and field HXS (Hart Index Shift) for
determining target addresses for MSIs is described later, in
Section~\ref{sec:AdvPLIC-MSIAddrs}.

Fields High Base PPN, Low Base PPN, and HXS are all {\WARL}.
An implementation is free to choose what values are supported.
Typically, some bits are writable while others are read-only constants.
In the extreme, the values of all three fields may be entirely
constant, fixed by the implementation.

\begin{commentary}
For most \mbox{32-bit} {\RISCV} systems, and for some \mbox{64-bit}
systems as well, it is expected that \z{mmsiaddrcfgh} will be a
constant zero.
\end{commentary}

If bit~L in \z{mmsiaddrcfg} is set to one, \z{mmsiaddrcfg} and
\z{mmsiaddrcfgh} are \emph{locked}, and writes to the registers are
ignored, making the registers effectively read-only.
When L~=~1, the other fields in \z{mmsiaddrcfg} and \z{mmsiaddrcfgh}
may optionally read as zeros.
In that case, if these other fields were given nonzero values
when L was first set, their values are retained internally by the
PLIC but become no longer visible by reading \z{mmsiaddrcfg} and
\z{mmsiaddrcfgh}.

At system reset, L~is initialized to either zero or one, whichever is
deemed appropriate for the specific PLIC implementation.
If reset initializes L to one, either the Base PPN and HXS values are
hardwired by the PLIC to constants, or the PLIC has a different means,
outside of this standard, for determining the addresses of outgoing MSI
writes.
In the latter case, the other fields in \z{mmsiaddrcfg} and
\z{mmsiaddrcfgh} may read as zeros, so registers \z{mmsiaddrcfg} and
\z{mmsiaddrcfgh} have only read-only values \z{0x00000001} and zero
respectively.
Any time \z{mmsiaddrcfg} or \z{mmsiaddrcfgh} has a different value
(not \z{0x00000001} or zero respectively), the addresses for
outgoing MSI writes directed to machine level must be derivable
from the visible values of Base PPN and HXS, as specified in
Section~\ref{sec:AdvPLIC-MSIAddrs}.

@@

PLICs that actually hardwire the IMSIC addresses internally
can implement these registers simply as read-only with values
\z{0x00000001} and zero.
Or, if the IMSIC addresses must be configured by software but
the formula is too complex for registers \z{mmsiaddrcfg} and
\z{mmsiaddrcfgh} to handle, again the registers can be implemented
simply as read-only with values \z{0x00000001} and zero, and a
separate, custom mechanism supplied for configuring the IMSIC
addresses.
\end{commentary}

@@

\textbf{%
The parameters used to configure MSI addresses will likely change from
what is currently described here for registers \z{smsiaddrcfg} and
\z{smsiaddrcfgh}.%
}
\vspace{3mm}

@@

When implemented, \z{smsiaddrcfg} has this format:\nopagebreak
\begin{displayLinesTable}[l@{\quad}l]
bits 31:12 & Low Base PPN (\WARL) \\
bits 10:8  & HXS (\WARL) \\
bit 0      & L \\
\end{displayLinesTable}
All other bits of \z{smsiaddrcfg} are reserved and read as zeros.
Register \z{smsiaddrcfgh} has this format:\nopagebreak
\begin{displayLinesTable}[l@{\ \quad}l]
bits 31:0 & High Base PPN (\WARL) \\
\end{displayLinesTable}

Fields High Base PPN from \z{smsiaddrcfgh} and Low Base PPN from
\z{smsiaddrcfg} concatenate to form a \mbox{52-bit} Base PPN
(Physical Page Number).
The use of this value and field HXS (Hart Index Shift) for
determining target addresses for MSIs is described later, in
Section~\ref{sec:AdvPLIC-MSIAddrs}.

Fields High Base PPN, Low Base PPN, and HXS are all {\WARL}.
An implementation is free to choose what values are supported, just as
for \z{mmsiaddrcfg} and \z{mmsiaddrcfgh}.

If bit~L in \z{smsiaddrcfg} is set to one, \z{smsiaddrcfg} and
\z{smsiaddrcfgh} are \emph{locked}, read-only.
When L~=~1, the other fields in \z{smsiaddrcfg} and \z{smsiaddrcfgh}
may optionally read as zeros.
In that case, if these other fields were given nonzero values
when L was first set, their values are retained internally by the
PLIC but become no longer visible by reading \z{smsiaddrcfg} and
\z{smsiaddrcfgh}.

At system reset, L~is initialized to either zero or one, whichever is
deemed appropriate for the specific PLIC implementation.
If reset initializes L to one, either the Base PPN and HXS values are
hardwired by the PLIC to constants, or the PLIC has a different means,
outside of this standard, for determining the addresses of outgoing MSI
writes for supervisor-level child domains of this domain.
In the latter case, the other fields in \z{smsiaddrcfg} and
\z{smsiaddrcfgh} may read as zeros, so registers \z{smsiaddrcfg} and
\z{smsiaddrcfgh} have only read-only values \z{0x00000001} and zero
respectively.
Any time \z{smsiaddrcfg} or \z{smsiaddrcfgh} has a different value
(not \z{0x00000001} or zero respectively), the addresses for outgoing
MSI writes from supervisor-level child domains must be derivable
from the visible values of Base PPN and HXS, as specified in
Section~\ref{sec:AdvPLIC-MSIAddrs}.

@@

\textbf{%
The optional configuration of MSI addresses using registers
\z{mmsiaddrcfg}, \z{mmsiaddrcfgh}, \z{smsiaddrcfg}, and
\z{smsiaddrcfgh} will likely change from what is currently described in
this subsection.%
}
\vspace{3mm}

@@

For a machine-level interrupt domain, if MSI target addresses are
determined by \z{mmsiaddrcfg} and \z{mmsiaddrcfgh}, then the address
for an outgoing MSI for interrupt source~$i$ is computed from those
registers and from the Hart Index field of register \z{target[$i$]} as
follows:
\begin{displayLinesTable}
$\mbox{MSI address} =
  \bigl(\mbox{Base PPN \z{|} (Hart Index\z{<<}HXS)}\bigr)\mbox{\z{<<}12}$
\end{displayLinesTable}
The vertical bar (\z{|}) here represents bitwise logical OR, and
\z{<<}$\,k$ represents shifting left by $k$ bits.

For a supervisor-level domain, if MSI target addresses are determined
by the parent domain's \z{smsiaddrcfg} and \z{smsiaddrcfgh} registers,
then to construct the address for an outgoing MSI for interrupt
source~$i$, the Hart Index from register \z{target[$i$]} must first be
converted into the index that the machine-level parent domain uses for
the same hart.
(Typically, these numbers are the same, but they may not be.)
The address for the MSI is then computed using this machine-level hart
index, the parent domain's \z{smsiaddrcfg} and \z{smsiaddrcfgh}, and
the Guest Index value from \z{target[$i$]}:
\begin{displayLinesTable}
$\mbox{MSI address} =
  \bigl(\mbox{Base PPN \z{|} (machine-level hart index\z{<<}HXS)
      \z{|} Guest Index}\bigr)
    \mbox{\z{<<}12}$
\end{displayLinesTable}

>>>>

@@

When implemented, \z{mmsiaddrcfg} has this format:\nopagebreak
\begin{displayLinesTable}[l@{\quad}l]
bits 31:0 & Low Base PPN (\WARL) \\
\end{displayLinesTable}
and \z{mmsiaddrcfgh} has this format:\nopagebreak
\begin{displayLinesTable}[l@{\quad}l]
bit 31     & L \\
bits 28:24 & HHXS (\WARL) \\
bits 22:20 & LHXS (\WARL) \\
bits 18:16 & HHXW (\WARL) \\
bits 15:12 & LHXW (\WARL) \\
bits 11:0  & High Base PPN (\WARL) \\
\end{displayLinesTable}
All other bits of \z{mmsiaddrcfgh} are reserved and read as zeros.

Fields High Base PPN from \z{mmsiaddrcfgh} and Low Base PPN from
\z{mmsiaddrcfg} concatenate to form a \mbox{44-bit} Base PPN
(Physical Page Number).
The use of this value and fields HHXS (High Hart Index Shift),
LHXS (Low Hart Index Shift), HHXW (High Hart Index Width),
and LHXW (Low Hart Index Width) for
determining target addresses for MSIs is described later, in
Section~\ref{sec:AdvPLIC-MSIAddrs}.

All fields other than L are {\WARL}.
An implementation is free to choose what values are supported.
Typically, some bits are writable while others are read-only constants.
In the extreme, the values of all fields may be entirely
constant, fixed by the implementation.

If bit~L in \z{mmsiaddrcfgh} is set to one, \z{mmsiaddrcfg} and
\z{mmsiaddrcfgh} are \emph{locked}, and writes to the registers are
ignored, making the registers effectively read-only.
When L~=~1, the other fields in \z{mmsiaddrcfg} and \z{mmsiaddrcfgh}
may optionally read as zeros.
In that case, if these other fields were given nonzero values
when L was first set, their values are retained internally by the
PLIC but become no longer visible by reading \z{mmsiaddrcfg} and
\z{mmsiaddrcfgh}.

At system reset, L~is initialized to either zero or one, whichever is
deemed appropriate for the specific PLIC implementation.
If reset initializes L to one, either the other fields are
hardwired by the PLIC to constants, or the PLIC has a different means,
outside of this standard, for determining the addresses of outgoing MSI
writes.
In the latter case, the other fields in \z{mmsiaddrcfg} and
\z{mmsiaddrcfgh} may read as zeros, so registers \z{mmsiaddrcfg} and
\z{mmsiaddrcfgh} have only read-only values zero and \z{0x80000000}
respectively.
Any time \z{mmsiaddrcfg} or \z{mmsiaddrcfgh} has a different value
(not zero or \z{0x80000000} respectively), the addresses for
outgoing MSI writes directed to machine level must be derivable
from the visible values of these registers, as specified in
Section~\ref{sec:AdvPLIC-MSIAddrs}.

@@

PLICs that actually hardwire the IMSIC addresses internally
can implement these registers simply as read-only with values zero
and \z{0x80000000}.
Or, if the IMSIC addresses must be configured by software but
the formula is too complex for registers \z{mmsiaddrcfg} and
\z{mmsiaddrcfgh} to handle, again the registers can be implemented
simply as read-only with values zero and \z{8x00000000}, and a
separate, custom mechanism supplied for configuring the IMSIC
addresses.
\end{commentary}

@@

@@

When implemented, \z{smsiaddrcfg} has this format:\nopagebreak
\begin{displayLinesTable}[l@{\quad}l]
bits 31:0 & Low Base PPN (\WARL) \\
\end{displayLinesTable}
and \z{smsiaddrcfgh} has this format:\nopagebreak
\begin{displayLinesTable}[l@{\quad}l]
bit 31     & L \\
bits 28:24 & HHXS (\WARL) \\
bits 22:20 & LHXS (\WARL) \\
bits 18:16 & HHXW (\WARL) \\
bits 15:12 & LHXW (\WARL) \\
bits 11:0  & High Base PPN (\WARL) \\
\end{displayLinesTable}
All other bits of \z{smsiaddrcfgh} are reserved and read as zeros.

Fields High Base PPN from \z{smsiaddrcfgh} and Low Base PPN from
\z{smsiaddrcfg} concatenate to form a \mbox{44-bit} Base PPN
(Physical Page Number).
The use of this value and fields HHXS (High Hart Index Shift),
LHXS (Low Hart Index Shift), HHXW (High Hart Index Width),
and LHXW (Low Hart Index Width) for
determining target addresses for MSIs is described later, in
Section~\ref{sec:AdvPLIC-MSIAddrs}.

All fields other than L are {\WARL}.
An implementation is free to choose what values are supported, just as
for \z{mmsiaddrcfg} and \z{mmsiaddrcfgh}.

If bit~L in \z{smsiaddrcfgh} is set to one, \z{smsiaddrcfg} and
\z{smsiaddrcfgh} are \emph{locked}, read-only.
When L~=~1, the other fields in \z{smsiaddrcfg} and \z{smsiaddrcfgh}
may optionally read as zeros.
In that case, if these other fields were given nonzero values
when L was first set, their values are retained internally by the
PLIC but become no longer visible by reading \z{smsiaddrcfg} and
\z{smsiaddrcfgh}.

At system reset, L~is initialized to either zero or one, whichever is
deemed appropriate for the specific PLIC implementation.
If reset initializes L to one, either the other fields are
hardwired by the PLIC to constants, or the PLIC has a different means,
outside of this standard, for determining the addresses of outgoing MSI
writes for supervisor-level child domains of this domain.
In the latter case, the other fields in \z{smsiaddrcfg} and
\z{smsiaddrcfgh} may read as zeros, so registers \z{smsiaddrcfg} and
\z{smsiaddrcfgh} have only read-only values zero and \z{0x80000000}
respectively.
Any time \z{smsiaddrcfg} or \z{smsiaddrcfgh} has a different value
(not zero or \z{0x80000000} respectively), the addresses for outgoing
MSI writes from supervisor-level child domains must be derivable
from the visible values of these registers, as specified in
Section~\ref{sec:AdvPLIC-MSIAddrs}.

@@

@@

For a machine-level interrupt domain, if MSI target addresses are
determined by \z{mmsiaddrcfg} and \z{mmsiaddrcfgh}, then the address
for an outgoing MSI for interrupt source~$i$ is computed from those
registers and from the Hart Index field of register \z{target[$i$]} as
follows:
\begin{displayLinesTable}
$g =
  (\mbox{Hart Index\z{>>}LHXW})\mbox{ \z{\&} }(\mbox{2}^{\rm HHXW}-\mbox{1})$\\
$h = \mbox{Hart Index \z{\&} }(\mbox{2}^{\rm LHXW}-\mbox{1})$\\
$\mbox{MSI address} =
  \bigl(\,
    \mbox{Base PPN \z{|} }(g\mbox{\z{<<}}(\mbox{HHXS}+\mbox{12}))
      \mbox{ \z{|} }(h\mbox{\z{<<}LHXS})
  \,\bigr)\mbox{\z{<<}12}$
\end{displayLinesTable}
Here, \z{<<}$\,k$ and \z{>>}$\,k$ represent shifting left and right by
$k$ bits, an ampersand (\z{\&}) represents bitwise logical AND, and a
vertical bar (\z{|}) represents bitwise logical OR.
Assuming the recommendations of
Section~\ref{sec:IMSIC-systemMemRegions} are followed for the
arrangement of IMSIC interrupt files in the machine's address space,
value $g$ is intended to be the number of a hart group (always zero
if HHXW =~0), while $h$ is the number of the target hart within that
group.
Represented in the terms of Section~\ref{sec:IMSIC-systemMemRegions},
HHXW =~$j$, LHXW =~$k$, HHXS = ${E-24}$, LHXS = ${C-12}$, and
Base~PPN = $A$\z{>>}12.

For a supervisor-level domain, if MSI target addresses are determined
by the parent domain's \z{smsiaddrcfg} and \z{smsiaddrcfgh} registers,
then to construct the address for an outgoing MSI for interrupt
source~$i$, the Hart Index from register \z{target[$i$]} must first be
converted into the index that the machine-level parent domain uses for
the same hart.
(Typically, these numbers are the same, but they may not be.)
The address for the MSI is then computed using this machine-level hart
index, the parent domain's \z{smsiaddrcfg} and \z{smsiaddrcfgh}, and
the Guest Index value from \z{target[$i$]}:
\begin{displayLinesTable}
$g =
  (\mbox{machine-level hart index\z{>>}LHXW})
    \mbox{ \z{\&} }(\mbox{2}^{\rm HHXW}-\mbox{1})$\\
$h = \mbox{machine-level hart index \z{\&} }(\mbox{2}^{\rm LHXW}-\mbox{1})$\\
$\mbox{MSI address} =
  \bigl(\,
    \mbox{Base PPN \z{|} }(g\mbox{\z{<<}}(\mbox{HHXS}+\mbox{12}))
      \mbox{ \z{|} }(h\mbox{\z{<<}LHXS})\mbox{ \z{|} Guest Index}
  \,\bigr)\mbox{\z{<<}12}$
\end{displayLinesTable}

Represented in the terms of Section~\ref{sec:IMSIC-systemMemRegions},
HHXW =~$j$, LHXW =~$k$, HHXS = ${E-24}$, LHXS = ${D-12}$, and
Base~PPN = $B$\z{>>}12.

### 2021-06-09 doc/src/IMSIC.tex
Clarify addresses of IMSIC interrupt files, and support major subdivision of address space.
Add recommendations for how to arrange the memory regions of IMSIC
interrupt files in a machine's address space.  Also, modify an
Advanced PLIC's MSI address configuration registers (mmsiaddrcfg,
mmsiaddrcfgh, smsiaddrcfg, smsiaddrcfgh) to support the subdivision
of the machine's address space so that different groups of harts are
constrained to separate parts of the address space (usually because
the harts are on separate chips in a "multi-socket" configuration).

>>>>

%-----------------------------------------------------------------------
\section{Arrangement of the memory regions of multiple interrupt files}
\label{sec:IMSIC-systemMemRegions}

Each interrupt file that an IMSIC implements has its own memory
region as described in the previous section, occupying exactly one
\mbox{4-KiB} page of machine address space.
When practical, the memory pages of the machine-level interrupt files
of all IMSICs should be located together in one part of the address
space, and the memory pages of all supervisor-level and guest interrupt
files should similarly be located together in another part of the
address space, according to the rules below.

\begin{commentary}
The main reason for separating the machine-level interrupt files
from the other interrupt files in the address space is so harts that
implement physical memory protection (PMP) can grant supervisor-level
access to all supervisor-level and guest interrupt files using only a
single PMP table entry.
If the memory pages for machine-level interrupt files are instead
interleaved with those of lower-privilege interrupt files, the number
of PMP table entries needed for granting supervisor-level access to all
non-machine-level interrupt files could equal the number of harts in
the system.
\end{commentary}

If a machine's construction dictates that harts be subdivided into
groups, with each group relegated to its own portion of the address
space, then the best that can be achieved is to locate together the
machine-level interrupt files of each group of harts separately, and
likewise locate together the supervisor-level and guest interrupt files
of each group of harts separately.
This situation is further addressed later below.

\begin{commentary}
A system may divide harts into groups in the address space because each
group exists on a separate chip (or chiplet in a multi-chip module),
and weaving together the address spaces of the multiple chips is
impractical.
In that case, granting supervisor-level access to all non-machine-level
interrupt files takes one PMP table entry per group.
\end{commentary}

For the purpose of locating the memory pages of interrupt files in the
address space, assume each hart (or each hart within a group) has a
unique hart number that may or may not be related to the unique hart
identifiers (``hart IDs'') that the RISC-V Privileged Architecture
assigns to harts.
For convenient addressing, the memory pages of all machine-level
interrupt files (or all those of a single group of harts) should be
arranged so that the address of the machine-level interrupt file for
hart number~$h$ is given by the formula ${A+h\times\mbox{2}^{C}}$ for
some integer constants $A$ and~$C$.
If the largest hart number is $h_{\rm max}$, let
${k = \lceil\log_{2}(h_{\rm max}+\mbox{1})\rceil}$, the number of bits
needed to represent any hart number.
Then the base address $A$ should be aligned to a $\mbox{2}^{k+C}$
address boundary, so ${A+h\times\mbox{2}^{C}}$ always equals
$A$~\z{|}~${(h\times\mbox{2}^{C})}$, where the vertical bar (\z{|})
represents bitwise logical OR.

The smallest that $C$ can be is~12, with $\mbox{2}^{C}$ being the size
of one \mbox{4-KiB} page.
If ${C > 12}$, the start of the memory page for each machine-level
interrupt file is aligned not just to a \mbox{4-KiB} page but to a
stricter $\mbox{2}^{C}$ address boundary.
Within the ${\mbox{2}^{k+C}}$-size address range $A$ through
${A+\mbox{2}^{k+C}-\mbox{1}}$, every \mbox{4-KiB} page that is not
occupied by a machine-level interrupt file should be filled with
\mbox{32-bit} words of read-only zeros, such that any read of an
aligned word returns zero and any write to an aligned word is ignored.

The memory pages of all supervisor-level interrupt files (or all those
of a single group of harts) should similarly be arranged so that the
address of the supervisor-level interrupt file for hart number~$h$
is ${B+h\times\mbox{2}^{D}}$ for some integer constants $B$ and~$D$,
with the base address $B$ being aligned to a $\mbox{2}^{k+D}$ address
boundary.

If an IMSIC implements guest interrupt files, the memory pages for the
IMSIC's supervisor-level interrupt file and for its guest interrupt
files are contiguous, starting with the supervisor-level interrupt
file at the lowest address and followed by the guest interrupt files,
ordered by guest interrupt number.
Schematically, the memory pages are ordered contiguously as
\begin{displayLinesTable}
S,\, $\mbox{G}_{1}$, $\mbox{G}_{2}$, $\mbox{G}_{3}$, \dots
\end{displayLinesTable}
where S~is the page for the supervisor-level interrupt file and each
$\mbox{G}_{i}$ is the page for the interrupt file of guest interrupt
number~$i$.
Consequently, the smallest that constant $D$ can be is
${\lceil\log_{\rm 2}(\mbox{maximum GEILEN}+\mbox{1})\rceil}+12$,
recalling that GEILEN for each IMSIC is the number of guest interrupt
files the IMSIC implements.

Within the ${\mbox{2}^{k+D}}$-size address range $B$ through
${B+\mbox{2}^{k+D}-\mbox{1}}$, every \mbox{4-KiB} page that is not
occupied by an interrupt file (supervisor-level or guest) should be
filled with \mbox{32-bit} words of read-only zeros.

When a system divides harts into groups, each in its own
separate portion of the address space, the memory page
addresses of interrupt files are expected to follow the formulas
${g\times\mbox{2}^{E}}+A+{h\times\mbox{2}^{C}}$ for machine-level
interrupt files, and ${g\times\mbox{2}^{E}}+B+{h\times\mbox{2}^{D}}$
for supervisor-level interrupt files, with $g$ being a
\emph{group number}, $h$ being a hart number relative to the group,
and $E$ being another integer constant $\geq$~${k+\max(C,D)}$ but
usually significantly larger.
If the largest group number is $g_{\rm max}$, let
${j = \lceil\log_{2}(g_{\rm max}+\mbox{1})\rceil}$, the number of bits
needed to represent any group number.
Besides being multiples of $\mbox{2}^{k+C}$ and $\mbox{2}^{k+D}$
respectively, $A$ and $B$ should be chosen so
\begin{displayLinesTable}[lll]
$\left((\mbox{2}^{j}-\mbox{1})\times\mbox{2}^{E}\right)$ \z{\&} $A \,=\, 0$ &
  and &
  $\left((\mbox{2}^{j}-\mbox{1})\times\mbox{2}^{E}\right)$ \z{\&} $B \,=\, 0$
\end{displayLinesTable}
where the ampersands (\z{\&}) represent bitwise logical AND.
This ensures that
\begin{displayLinesTable}[lcll]
$g\times\mbox{2}^{E}+A+h\times\mbox{2}^{C}$ & always equals &
  $(g\times\mbox{2}^{E})$ \z{|} $A$ \z{|} $(h\times\mbox{2}^{C})$, & and\\
$g\times\mbox{2}^{E}+B+h\times\mbox{2}^{D}$ & always equals &
  $(g\times\mbox{2}^{E})$ \z{|} $B$ \z{|} $(h\times\mbox{2}^{D})$.\\
\end{displayLinesTable}

The requirements for read-only-zero pages apply only within each group.
That is, for each group~$g$, within the address ranges
\begin{displayLinesTable}[lcll]
$g\times\mbox{2}^{E}+A$ & through &
  $g\times\mbox{2}^{E}+A+\mbox{2}^{k+C}-\mbox{1}$, &
  and\\
$g\times\mbox{2}^{E}+B$ & through &
  $g\times\mbox{2}^{E}+B+\mbox{2}^{k+D}-\mbox{1}$,\\
\end{displayLinesTable}
pages not occupied by an interrupt file should be read-only zeros.

See also Section \ref{sec:AdvPLIC-MSIAddrs} for the default algorithms
an Advanced PLIC may use to determine the destination addresses of
outgoing MSIs, which should be the addresses of IMSIC interrupt files.
